---
title: "Univariate linear regression"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(kableExtra)
```

Univariate linear regression to see if there is significant pathways from the humann2 normalized data

## Load pheno and count data 

```{r}
final_pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  dplyr::rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  dplyr::rename(pw = names(.)[1])  %>% 
  # remove the things after the : so that its easier to see on the plot(the pathway name)
  mutate(pw = str_replace(pw, ':.+$','')) 
```

##### Reorder the column and make sure the sample names in the pheno table align with the sample name in the count table

```{r}
pw <- bind_cols(pw[,1],pw[,final_pheno$sampleid])

all.equal(colnames(pw)[2:ncol(pw)], final_pheno$sampleid)
```

## filter to have each pathway with at least 50% of nonzero (zero <= 50%)

```{r}
# filter to have each pathway with at least 50% of nonzero (zero <= 50%)
pw_fil <- pw %>% 
  column_to_rownames('pw') %>% 
  as.matrix() 

num_zero_thre <- floor(ncol(pw_fil) * 0.5)

pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]

```

##### How many pathways left

```{r}
nrow(pw_fil)
```

## Select the most abundant 10 pathways and log2 transform the data

```{r}
# select the most abundant N pathways and log2 transform the data

N <- 10
topN_pw <- sort(rowSums(pw_fil), decreasing = T)[1:N]
pw_Fil <- pw_fil[names(topN_pw),]

pw_Fil <- log2(pw_Fil)

pw_Fil %>% 
  kable(caption = 'log2 transformed count table with top 10 most abundant pathways') %>% 
  kable_styling(full_width = F, position = 'left')
```

## Linear regression for each pathway with Toxicity and CR as response respectively, followed by FDR correction

```{r}
# do a linear regression of each rows and then do a FDR correction later
# the pheno in the same order as the current sample names 
phe_tox <- final_pheno$Toxicity
phe_CR <- final_pheno$CR

res_tox <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_tox))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))


res_CR <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_CR))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))

res_tox %>% 
  kable(caption = 'Response to Toxicity') %>% 
  kable_styling(full_width = F, position = 'left')

res_CR %>% 
  kable(caption = 'Response to CR') %>% 
  kable_styling(full_width = F, position = 'left')
```

## How many pathways are significant with FDR < 0.1

```{r}
# Toxicity
res_tox %>% 
  filter(padj < 0.1) %>% 
  nrow

# CR
res_CR %>% 
  filter(padj < 0.1) %>% 
  nrow

```


##### I have tested: no sig pathways even select the top 200 most abundant ones.