---
title: "16s genus relab and phenotype"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

Enteroccous relative abundance v CR D100

Enteroccous relative abundance v Toxicity

Toxicity grade (both crs and icans) in light of the relative abundance of bacteroidetes or enterococcus

## Enteroccous relative abundance v CR D100

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')
```

```{r}
# get the relative abundance of bacteroidetes or enterococcus for those samples
target <- genus %>% 
  filter(genus %in% c('Enterococcus','Bacteroides')) %>% 
  select(-center) %>% 
  spread('genus', 'relab') %>% 
  mutate(Bacteroides_log = log10(Bacteroides + 2*10^-6),
         Enterococcus_log = log10(Enterococcus + 2*10^-6)) 

joined <- meta %>% 
  full_join(target) %>% 
  mutate(center = if_else(center == 'P', 'Upenn', "MSK"))
```

```{r}
# a tally of the yes and no and unknown for the CR D100 to put in the title
joined %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  ggboxplot(x = 'cr_d100', y = 'Enterococcus_log', facet.by = 'center', add = 'jitter', color = 'cr_d100',
            xlab = 'CR Day 100', ylab = 'log10(Enterococcus relab)',
            title = 'CR D100 VS log10(Enterococcus relab)') +
  scale_color_manual(values = c('#ED0000','#00468B'))  +
  stat_compare_means(comparisons= list(c('yes', 'no')),
    label= "p.format",
    method= 'wilcox.test',
    exact=FALSE) +
  ggsave('../figs/amplicon/CR D100 VS log10(Enterococcus relab).pdf', width = 5, height = 5)
```

## Enteroccous relative abundance v Toxicity

```{r}
joined %>% 
  ggboxplot(x = 'toxicity', y = 'Enterococcus_log', facet.by = 'center', add = 'jitter', color = 'toxicity',
            xlab = 'toxicity', ylab = 'log10(Enterococcus relab)',
            title = 'toxicity VS log10(Enterococcus relab)') +
  scale_color_manual(values = c('#0099B4','#AD002A'))  +
   stat_compare_means(comparisons= list(c('yes', 'no')),
    label= "p.format",
    method= 'wilcox.test',
    exact=FALSE) +
  ggsave('../figs/amplicon/toxicity VS log10(Enterococcus relab).pdf', width = 5, height = 4.5)
```
## Toxicity grade (both crs and icans) VS the relative abundance of bacteroidetes or enterococcus

```{r}
joined %>% 
  count(center, icansgrade)

joined %>% 
  count(center, crsgrade)

# ignore the pending ones
df <- joined %>% 
  mutate(icansgrade = as.numeric(icansgrade)) %>% 
  filter(!is.na(icansgrade)) %>% 
  select(icansgrade, crsgrade, Bacteroides_log, Enterococcus_log, center)  %>% 
  gather('item', 'grade', icansgrade:crsgrade)
   
df %>% 
  ggboxplot(x = 'grade', y = 'Bacteroides_log',  add = 'jitter',
            xlab = 'Toxicity grade', ylab = 'log10(Bacteroides relab)',
            title = 'Toxicity grade VS log10(Bacteroides relab)') +
  facet_wrap(item ~ center, scales = 'free') +
  stat_compare_means(comparisons= list(c('0', '1'), c('0','2'), c('0','3')),label= "p.format", method= 'wilcox.test',exact=FALSE)  + 
  ggsave('../figs/amplicon/toxicity grade VS log10(Bacteroides relab).pdf', width = 7, height = 7)
```
```{r}
df %>% 
  ggboxplot(x = 'grade', y = 'Enterococcus_log',  add = 'jitter',
            xlab = 'Toxicity grade', ylab = 'log10(Enterococcus relab)',
            title = 'Toxicity grade VS log10(Enterococcus relab)') +
  facet_wrap(item ~ center, scales = 'free') +
  stat_compare_means(comparisons= list(c('0', '1'), c('0','2'), c('0','3')),label= "p.format", method= 'wilcox.test',exact=FALSE)  + 
  ggsave('../figs/amplicon/toxicity grade VS log10(Enterococcus relab).pdf', width = 7, height = 7)
```
 The problem is when they don't have all the comparison in a specific facet there is gonna be no comparison and p value gotta think about this 
